
[To-dos (contains research
questions)](https://docs.google.com/document/d/1vTpxr06BUZmbJGiVwWKPF25CAjFsIY7gwTVsOitZcGI/edit)

[Meetings con
Paolo](https://docs.google.com/document/d/1MMMNPR3i4zqfXu89s6NPBws8MVdwPz0GxWhJaZTk59o/edit)

# Initial set up

Set working directory
```{r setup}
knitr::opts_knit$set(root.dir = "C:/Users/benji/Desktop/Regrid")
```

Load packages
```{r}
library(sf)
library(tidyverse)
library(ineq)
library(tmap)
```

Set universal variables
```{r}
tmap_mode("view")
options(scipen = 999)
```

# Analysis

## Inequality

### Gini of land

Here, we assess inequality in the distribution of cropland. We do this for cropland parcels with a well match, and as a robustness check, but for all cropland parcels in the Central Valley. Again, we aggregate by subbasin so that nearby farms with the same owner are considered as such rather than separate entities.


Let's calculate the gini coefficients for land area, improved parcel value, land value, and agricultural value.

```{r}
cropland_parcels_filtered_owners = read.csv("Data/Parcels/Cropland parcels/cropland_parcels_filtered_owners.csv")
#cropland_parcels_normalised = read.csv("Data/Parcels/Cropland parcels/cropland_parcels_normalised.csv")
```


```{r}
paste("Gini for land area owned:", round(ineq(cropland_parcels_filtered_owners$parcel_area_ha), 3))
paste("Gini for improved value of land owned:", round(ineq(cropland_parcels_filtered_owners$improvval), 3))
paste("Gini for land value of land owned:", round(ineq(cropland_parcels_filtered_owners$landval), 3))
paste("Gini for agricultural value of land owned:", round(ineq(cropland_parcels_filtered_owners$agval), 3))

nrow(cropland_parcels_filtered_owners%>%filter(!is.na(parcel_area_ha)))
nrow(cropland_parcels_filtered_owners%>%filter(!is.na(improvval)))
nrow(cropland_parcels_filtered_owners%>%filter(!is.na(landval)))
nrow(cropland_parcels_filtered_owners%>%filter(!is.na(agval)))
```
```{r}
calculate_top_percent <- function(data, variable, top_percent) {
  # Sort data by the variable in descending order, ignoring NAs
  sorted_data <- data[order(-data[[variable]], na.last = NA), ]
  
  # Calculate the cumulative sum of the variable for the top X% of rows
  n_top <- ceiling(nrow(sorted_data) * top_percent / 100)
  top_sum <- sum(sorted_data[[variable]][1:n_top], na.rm = TRUE)
  
  # Calculate the total sum of the variable, removing NAs
  total_sum <- sum(data[[variable]], na.rm = TRUE)
  
  # Calculate the percentage contribution of the top X%
   percent_contribution <- signif((top_sum / total_sum) * 100, 3)
  return(percent_contribution)}


# Apply to each variable
top_percentages <- data.frame(
  variable = c("parcel_area_ha", "landval", "improvval", "agval"),
  top_1_percent = sapply(c("parcel_area_ha", "landval", "improvval", "agval"), 
                         function(x) calculate_top_percent(cropland_parcels_filtered_owners, x, top_percent = 1)),
  top_10_percent = sapply(c("parcel_area_ha", "landval", "improvval", "agval"), 
                          function(x) calculate_top_percent(cropland_parcels_filtered_owners, x, top_percent = 10)))

top_percentages
```


Gini is highest for ag value.

Let's compare this to the ginis when just looking at the subset of parcels with matched wells

```{r}
final_matches_joined_filtered_owners = read.csv("Data/Well-parcel matches/final_matches_joined_filtered_owners.csv")
```

```{r}
paste("Gini for land area owned:", round(ineq(final_matches_joined_filtered_owners$parcel_area_ha), 3))
paste("Gini for improved value of land owned:", round(ineq(final_matches_joined_filtered_owners$improved_val), 3))
paste("Gini for land value of land owned:", round(ineq(final_matches_joined_filtered_owners$land_val), 3))
paste("Gini for agricultural value of land owned:", round(ineq(final_matches_joined_filtered_owners$ag_val), 3))

nrow(final_matches_joined_filtered_owners%>%filter(!is.na(parcel_area_ha)))
nrow(final_matches_joined_filtered_owners%>%filter(!is.na(improved_val)))
nrow(final_matches_joined_filtered_owners%>%filter(!is.na(land_val)))
nrow(final_matches_joined_filtered_owners%>%filter(!is.na(ag_val)))
```

```{r}
top_percentages <- data.frame(
  variable = c("parcel_area_ha", "land_val", "improved_val", "ag_val"),
  top_1_percent = sapply(c("parcel_area_ha", "land_val", "improved_val", "ag_val"),
                         function(x) calculate_top_percent(final_matches_joined_filtered_owners, x, top_percent = 1)),
  top_10_percent = sapply(c("parcel_area_ha", "land_val", "improved_val", "ag_val"),
                          function(x) calculate_top_percent(final_matches_joined_filtered_owners, x, top_percent = 10)))

top_percentages
```


It varies quite a bit (sometimes higher, sometimes lower). I would probably trust all cropland parcels more? Ask Paolo.



### Gini of water

Here, we investigate the gini coefficient for well depth and well
capacity as metrics of groundwater access. We perform this analysis for
matched wells, and as a robustness check, for all wells marked for
irrigation use.

```{r}
final_matches_joined_filtered_owners = read.csv("Data/Well-parcel matches/final_matches_joined_filtered_owners.csv")
final_matches_joined_normalised = read.csv("Data/Well-parcel matches/final_matches_joined_normalised.csv")
```

```{r}
#only matched wells, filtering out public wells
paste("Gini for well depth:", round(ineq(final_matches_joined_filtered_owners$well_depth_ft), 3))
paste("Gini for well capacity:", round(ineq(final_matches_joined_filtered_owners$well_capacity_gpm), 3))

nrow(final_matches_joined_filtered_owners%>%filter(!is.na(well_capacity_gpm)))
nrow(final_matches_joined_filtered_owners%>%filter(!is.na(well_depth_ft)))


final_matches_joined_filtered_owners

#only matched wells filtering out public wells (normalised)
#ineq(final_matches_joined_normalised$depth_z_score - min(final_matches_joined_normalised$depth_z_score, na.rm = T))
#ineq(final_matches_joined_normalised$yield_z_score - min(final_matches_joined_normalised$yield_z_score, na.rm = T))
```

```{r}
wells = read.csv("Data/Wells/Processed/filtered_wells_with_section.gpkg")%>%
  filter(grepl("Irrigation", USGS.Water.Use.Category))
```


```{r}
#all wells
paste("Gini for well depth:", round(ineq(wells$Well.Depth..Feet.), 3))
paste("Gini for well depth:", round(ineq(wells$Well.Capacity..GPM.), 3))

nrow(wells%>%filter(!is.na(Well.Capacity..GPM.)))
nrow(wells%>%filter(!is.na(Well.Depth..Feet.)))
```

Well capacity appears to be more unequally distributed across Central
Valley well owners compared to well depth. This makes sense, as we would
expect well capacity (as the actual volume that can be pumped out) would
correlate more directly to groundwater access / ownership.

### Land-water

Now, let's explore relationships between land ownership and groundwater
access.

First, let's just make some scatter plots of land area vs metrics of
groundwater access (well depth and well yield). We do this

```{r}
options(scipen = 999)
final_matches_joined_normalised%>%
  ggplot(aes(area_z_score, depth_z_score))+
  geom_point(colour = "#410505")+
  scale_x_log10()+
  scale_y_log10()+
  labs(x = "Parcel Area Z-Score", y = "Well Depth Z-Score")+
  geom_smooth(method = "lm", linetype = "dashed", color = "coral3", fill = "coral1") +
  theme_bw()
final_matches_joined_filtered_owners%>%
  ggplot(aes(parcel_area_ha, well_depth_ft))+
  geom_point(colour = "#410505")+
  scale_x_log10()+
  scale_y_log10()+
  labs(x = "Parcel Area (ha)", y = "Well Depth (ft)")+
  geom_smooth(method = "lm", linetype = "dashed", color = "coral3", fill = "coral1") +
  theme_bw()

final_matches_joined_normalised%>%
  ggplot(aes(area_z_score, yield_z_score))+
  geom_point(colour = "midnightblue")+
  scale_x_log10()+
  scale_y_log10()+
  labs(x = "Parcel Area Z-Score", y = "Well Capacity Z-Score")+
  geom_smooth(method = "lm", linetype = "dashed", color = "skyblue4", fill = "skyblue2") +
  theme_bw()
final_matches_joined_filtered_owners%>%
  ggplot(aes(parcel_area_ha, well_capacity_gpm))+
  geom_point(colour = "midnightblue")+
  scale_x_log10()+
  scale_y_log10()+
  labs(x = "Parcel Area (ha)", y = "Well Capacity (gpm)")+
  geom_smooth(method = "lm", linetype = "dashed", color = "skyblue4", fill = "skyblue2") +
  theme_bw()
final_matches_joined_normalised%>%
  ggplot(aes(parcel_area_ha, yield_z_score))+
  geom_point(colour = "midnightblue")+
  scale_x_log10()+
  scale_y_log10()+
  labs(x = "Parcel Area (ha)", y = "Well Capacity (gpm)")+
  geom_smooth(method = "lm", linetype = "dashed", color = "skyblue4", fill = "skyblue2") +
  theme_bw()
```

The relationship looks relatively linear on the plots, which considering
the log x and log y axes, means that the true relationship between
normalised parcel area and normalised well metrics is non-linear (power
log, to be precise).

Now, let's visualise by percentile of land owned to make it easier to
pick out this trend.

Well Yield

```{r}
#Land percentile based on parcel area
final_matches_joined_filtered_owners%>%
  group_by(area_percentile)%>%
  summarise(mean = mean(well_capacity_gpm, na.rm = T))%>%
  ggplot(aes(area_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned", y = "Mean Well Capacity (GPM)")

final_matches_joined_normalised%>%
  group_by(area_percentile)%>%
  summarise(mean = mean(yield_z_score, na.rm = T))%>%
  ggplot(aes(area_percentile, mean))+
  geom_point()+
  #scale_x_discrete(breaks = seq(10, 100, by = 10))+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned", y = "Mean Well Capacity Z-Score")

#Land percentile based on parcel area normalised (z-score) by subbasin
final_matches_joined_normalised%>%
  group_by(area_percentile_z)%>%
  summarise(mean = mean(well_capacity_gpm, na.rm = T))%>%
  ggplot(aes(area_percentile_z, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned (Normalised)", y = "Mean Well Capacity (GPM)")
final_matches_joined_normalised%>%
  group_by(area_percentile_z)%>%
  summarise(mean = mean(yield_z_score, na.rm = T))%>%
  ggplot(aes(area_percentile_z, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned (Normalised)", y = "Mean Well Capacity Z-Score")
```

Here we clearly see the same trend: the curve is clearly non-linear,
indicating that those who own more cropland have disproportionately more
access to groundwater in terms of well depth and well capacity. This is
robust to normalising both well capacity and land area owned by subbasin
as well.

Well Depth

```{r}
final_matches_joined_normalised%>%
  group_by(area_percentile)%>%
  summarise(mean = mean(well_depth_ft, na.rm = T))%>%
  ggplot(aes(area_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned", y = "Mean Well Depth (ft)")
final_matches_joined_normalised%>%
  group_by(area_percentile)%>%
  summarise(mean = mean(depth_z_score, na.rm = T))%>%
  ggplot(aes(area_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned", y = "Mean Well Depth Z-Score")

final_matches_joined_normalised%>%
  group_by(area_percentile_z)%>%
  summarise(mean = mean(well_depth_ft, na.rm = T))%>%
  ggplot(aes(area_percentile_z, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned (Normalised)", y = "Mean Well Depth (ft)")
final_matches_joined_normalised%>%
  group_by(area_percentile_z)%>%
  summarise(mean = mean(depth_z_score, na.rm = T))%>%
  ggplot(aes(area_percentile_z, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned (Normalised) ", y = "Mean Well Depth Z-Score")
```

Lastly, let's see how other land ownership metrics, such as land value,
play out with these trends

Land value

```{r}
final_matches_joined_normalised%>%
  group_by(land_val_percentile)%>%
  summarise(mean = mean(depth_z_score, na.rm = T))%>%
  ggplot(aes(land_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Value Owned (Normalised)", y = "Mean Well Depth Z-Score")
final_matches_joined_filtered_owners%>%
  group_by(land_val_percentile)%>%
  summarise(mean = mean(well_depth_ft, na.rm = T))%>%
  ggplot(aes(land_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Value Owned", y = "Mean Well Depth Z-Score")


final_matches_joined_normalised%>%
  group_by(land_val_percentile)%>%
  summarise(mean = mean(yield_z_score, na.rm = T))%>%
  ggplot(aes(land_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Value Owned (Normalised)", y = "Mean Well Capacity Z-Score")
final_matches_joined_filtered_owners%>%
  group_by(land_val_percentile)%>%
  summarise(mean = mean(well_capacity_gpm, na.rm = T))%>%
  ggplot(aes(land_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Value Owned", y = "Mean Well Capacity Z-Score")
```

Improved value

```{r}
final_matches_joined_normalised%>%
  group_by(improved_val_percentile)%>%
  summarise(mean = mean(depth_z_score, na.rm = T))%>%
  ggplot(aes(improved_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Improved Land Value Owned (Normalised)", y = "Mean Well Depth Z-Score")
final_matches_joined_filtered_owners%>%
  group_by(improved_val_percentile)%>%
  summarise(mean = mean(well_depth_ft, na.rm = T))%>%
  ggplot(aes(improved_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Improved Land Value Owned", y = "Mean Well Depth (ft)")


final_matches_joined_normalised%>%
  group_by(improved_val_percentile)%>%
  summarise(mean = mean(yield_z_score, na.rm = T))%>%
  ggplot(aes(improved_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Improved Land Value Owned (Normalised)", y = "Mean Well Capacity Z-Score")
final_matches_joined_filtered_owners%>%
  group_by(improved_val_percentile)%>%
  summarise(mean = mean(well_capacity_gpm, na.rm = T))%>%
  ggplot(aes(improved_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Improved Land Value Owned", y = "Mean Well Capacity (gpm)")

```

Agricultural value

```{r}
final_matches_joined_normalised%>%
  group_by(ag_val_percentile)%>%
  summarise(mean = mean(depth_z_score, na.rm = T))%>%
  ggplot(aes(ag_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Agricultural Land Value Owned", y = "Mean Well Depth Z-Score")
final_matches_joined_filtered_owners%>%
  group_by(ag_val_percentile)%>%
  summarise(mean = mean(well_depth_ft, na.rm = T))%>%
  ggplot(aes(ag_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Agricultural Land Value Owned", y = "Mean Well Depth (ft)")


final_matches_joined_normalised%>%
  group_by(ag_val_percentile)%>%
  summarise(mean = mean(yield_z_score, na.rm = T))%>%
  ggplot(aes(ag_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Agricultural Land Value Owned", y = "Mean Well Capacity Z-Score")
final_matches_joined_filtered_owners%>%
  group_by(ag_val_percentile)%>%
  summarise(mean = mean(well_capacity_gpm, na.rm = T))%>%
  ggplot(aes(ag_val_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Agricultural Land Value Owned", y = "Mean Well Capacity (gpm)")
```

```{r}
final_matches_joined_normalised%>%
  group_by(area_percentile)%>%
  summarise(mean = mean(newest_well, na.rm = T))%>%
  ggplot(aes(area_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned", y = "Mean Newest Well Year")
final_matches_joined_normalised%>%
  group_by(area_percentile)%>%
  summarise(mean = mean(well_year_z_score, na.rm = T))%>%
  ggplot(aes(area_percentile, mean))+
  geom_point()+
  theme_bw()+
  labs(x = "Percentile of Land Area Owned", y = "Mean Newest Well Year Z-Score")
```

```{r}
total_yield = sum(final_matches_joined_filtered_owners%>%filter(!is.na(well_capacity_gpm))%>%pull(well_capacity_gpm), na.rm = T)
total_area = sum(final_matches_joined_filtered_owners%>%filter(!is.na(well_capacity_gpm))%>%pull(parcel_area_ha), na.rm = T)

final_matches_joined_filtered_owners %>%
  arrange(well_capacity_gpm)%>%  # Sort by well capacity
  filter(!is.na(well_capacity_gpm))%>%
  select(parcel_area_ha, well_capacity_gpm)%>%
  mutate(
    cum_land = cumsum(parcel_area_ha) / total_area ,
    cum_yield = cumsum(well_capacity_gpm) / total_yield)%>%
  arrange(parcel_area_ha)%>%
  ggplot(aes(x = cum_land, y = cum_yield)) +
    geom_point(color = "blue", size = 1.2) +   # Lorenz curve
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1.2) +  # Line of equality
    labs(
      title = "Lorenz Curve of Well Yield Owned by Land Area Owned",
      y = "Cumulative Share of Water Yield for Wells Owned",
      x = "Cumulative Share of Land Area Owned") +
    theme_minimal()


total_depth= sum(final_matches_joined_filtered_owners%>%filter(!is.na(well_depth_ft))%>%pull(well_depth_ft), na.rm = T)
total_area = sum(final_matches_joined_filtered_owners%>%filter(!is.na(well_depth_ft))%>%pull(parcel_area_ha), na.rm = T)

final_matches_joined_filtered_owners %>%
  arrange(well_depth_ft)%>%  # Sort by well depth
  filter(!is.na(well_depth_ft))%>%
  select(parcel_area_ha, well_depth_ft)%>%
  mutate(
    cum_land = cumsum(parcel_area_ha)/total_area,
    cum_depth = cumsum(well_depth_ft)/total_depth)%>%
  arrange(parcel_area_ha)%>%
  ggplot(aes(x = cum_land, y = cum_depth)) +
    geom_point(color = "blue", size = 1.2) +   # Lorenz curve
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1.2) +  # Line of equality
    labs(
      title = "Lorenz Curve of Well Depth Owned by Land Area Owned",
      y = "Cumulative Share of Water Depth for Wells Owned",
      x = "Cumulative Share of Land Area Owned") +
    theme_minimal()
```

For each farm, calculate water per unit area (well yield or well depth /
ha) and see what the distribution of that variable. plot area vs well
depth per area

```{r}
final_matches_joined_filtered_owners%>%
  mutate(depth_per_ha = well_depth_ft / parcel_area_ha,
         capacity_per_ha = well_capacity_gpm / parcel_area_ha)%>%
  group_by(area_percentile)%>%
  summarise(mean_depth = mean(depth_per_ha, na.rm = T))%>%
  ggplot(aes(area_percentile, mean_depth))+
    geom_point()+
    theme_bw()+
    labs(x = "Percentile of Land Area Owned", y = "Mean Well Depth per Hectare \n (ft per ha)")+
    scale_y_log10()

final_matches_joined_filtered_owners%>%
  mutate(depth_per_ha = well_depth_ft / parcel_area_ha,
         capacity_per_ha = well_capacity_gpm / parcel_area_ha)%>%
  group_by(area_percentile)%>%
  summarise(mean_yield = mean(capacity_per_ha, na.rm = T))%>%
  ggplot(aes(area_percentile, mean_yield))+
    geom_point()+
    theme_bw()+
    labs(x = "Percentile of Land Area Owned", y = "Mean Well Capacity per Hectare \n (GPM per ha)")+
    scale_y_log10()
```

Do you see more value per unit land area for parcels that have more gw
water access (well depth or well yield)

```{r}
final_matches_joined_filtered_owners%>%
  mutate(depth_percentile = ntile(well_depth_ft, 100), value_per_ha = land_val/parcel_area_ha)%>%
  group_by(depth_percentile)%>%
  reframe(value_per_ha = mean(value_per_ha, na.rm = T))%>%
  ggplot(aes(depth_percentile, value_per_ha))+
    geom_point()+
    theme_bw()+
    labs(x = "Percentile of Well Depth Ownership", y = "Land Value per ha ($/ha)")

final_matches_joined_filtered_owners%>%
  mutate(depth_percentile = ntile(well_depth_ft, 100),value_per_ha = improved_val/parcel_area_ha)%>%
  group_by(depth_percentile)%>%
  reframe(value_per_ha = mean(value_per_ha, na.rm = T))%>%
  ggplot(aes(depth_percentile, value_per_ha))+
    geom_point()

final_matches_joined_filtered_owners%>%
  mutate(depth_percentile = ntile(well_depth_ft, 100),value_per_ha = ag_val/parcel_area_ha)%>%
  group_by(depth_percentile)%>%
  reframe(value_per_ha = mean(value_per_ha, na.rm = T))%>%
  ggplot(aes(depth_percentile, value_per_ha))+
    geom_point()
```

```{r}
final_matches_joined_filtered_owners%>%
  mutate(yield_percentile = ntile(well_capacity_gpm, 100), value_per_ha = land_val/parcel_area_ha)%>%
  group_by(yield_percentile)%>%
  reframe(value_per_ha = mean(value_per_ha, na.rm = T))%>%
  ggplot(aes(yield_percentile, value_per_ha))+
    geom_point()+
    theme_bw()+
    labs(x = "Percentile of Well Capacity Ownership", y = "Land Value per ha ($/ha)")
  

final_matches_joined_filtered_owners%>%
  mutate(yield_percentile = ntile(well_capacity_gpm, 100),value_per_ha = improved_val/parcel_area_ha)%>%
  group_by(yield_percentile)%>%
  reframe(value_per_ha = mean(value_per_ha, na.rm = T))%>%
  ggplot(aes(yield_percentile, value_per_ha))+
    geom_point()
final_matches_joined_filtered_owners%>%
  mutate(yield_percentile = ntile(well_capacity_gpm, 100),value_per_ha = ag_val/parcel_area_ha)%>%
  group_by(yield_percentile)%>%
  reframe(value_per_ha = mean(value_per_ha, na.rm = T))%>%
  ggplot(aes(yield_percentile, value_per_ha))+
    geom_point()
```

## More descriptive stuff

### Precipitation and land size

Use chirps back to 1981

1.  Get mean annual precipitation by parcel. Relationship between precip
    and parcel size / parcel value
2.  By GW basin (or other boundary), relationship between mean
    precipitation and land inequality
3.  By GW basin (or other boundary), relationship between mean
    precipitation and mean land size/value

```{r}
precip = rast("Precipitation data/chirps-v2.0.annual.nc")

parcel_precip_means <- exact_extract(crop(precip, ext(master_cropland_parcels)), master_cropland_parcels, 'mean', force_df = TRUE)%>%
  rename_with(~ as.character(1981:2023))%>%
  mutate(mean_annual_precip = rowMeans(across(everything())))%>%
  pull(mean_annual_precip)

master_cropland_parcels_df = master_cropland_parcels%>%
  mutate(mean_annual_precip = parcel_precip_means)%>%
  data.frame()

master_cropland_parcels_df

summary(lm(data = master_cropland_parcels_df, parcel_area_ha ~ mean_annual_precip))
```

```{r}
subbasin_precip_means <- exact_extract(crop(precip, ext(gw_basins)), gw_basins, 'mean', force_df = TRUE)%>%
  rename_with(~ as.character(1981:2023))%>%
  mutate(mean_annual_precip = rowMeans(across(everything())))%>%
  pull(mean_annual_precip)
gw_basins = gw_basins%>%
  mutate(mean_annual_precip = subbasin_precip_means)%>%
  filter(basin_number %in% c("5-022", "5-021"))

stats_by_subbasin = final_matches_joined_filtered_owners%>%
  group_by(subbasin_name)%>%
  reframe(
          median_parcel_area = median(parcel_area_ha, na.rm = T),
          gini_parcel_area = ineq(parcel_area_ha, na.rm = T),
          median_well_capacity = median(well_capacity_gpm, na.rm = T),
          median_well_depth = median(well_depth_ft, na.rm = T),
          median_well_depth = median(well_depth_ft, na.rm = T),
          gini_well_capacity = ineq(well_capacity_gpm, na.rm = T),
          gini_well_depth = ineq(well_depth_ft, na.rm = T),
          gini_land_value = ineq(land_val, na.rm =  T),
          gini_ag_value = ineq(ag_val, na.rm = T),
          gini_improved_value = ineq(improved_val, na.rm = T),
          n = n())%>%
  filter(n > 10)

gw_basins_full = merge(gw_basins%>%select(subbasin_number, subbasin_name, mean_annual_precip), stats_by_subbasin)
```


```{r}
library(classInt)
round_to_5 <- function(x) {round(x / 5) * 5}

tmap_mode("view")
tmap_mode("plot")

#precip
tm_shape(gw_basins) +
  tm_polygons(col = "mean_annual_precip", palette = "Blues", title = "Mean Annual Precipitation (mm)") 


#parcel area
jenks_breaks <- round_to_5(classIntervals(gw_basins_full$median_parcel_area, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons("median_parcel_area", breaks = jenks_breaks, palette = "Reds", title = "Median Parcel Area (ha)") 

jenks_breaks <- round_to_5(classIntervals(gw_basins_full$gini_parcel_area, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "gini_parcel_area", palette = "Blues", title = "Gini Coefficient of Parcel Area") 


#land value
jenks_breaks <- round_to_5(classIntervals(gw_basins_full$gini_land_value, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "gini_land_value", palette = "Purples", title = "Gini Coefficient of Land Value") 

jenks_breaks <- round_to_5(classIntervals(gw_basins_full$gini_improved_value, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "gini_improved_value", palette = "Purples", title = "Gini Coefficient of Improved Land Value")

#well capacity
jenks_breaks <- round_to_5(classIntervals(gw_basins_full$median_well_capacity, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "median_well_capacity", palette = "Blues", title = "Median Well Capacity (GPM)") 
jenks_breaks <- round_to_5(classIntervals(gw_basins_full$gini_well_capacity, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "gini_well_capacity", palette = "Blues", title = "Gini Coefficient of Well Capacity") 

#well depth
jenks_breaks <- round_to_5(classIntervals(gw_basins_full$median_well_depth, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "median_well_depth", palette = "RdPu", title = "Median Well Depth (ft)") 
jenks_breaks <- round_to_5(classIntervals(gw_basins_full$gini_well_depth, n = 5, style = "jenks")$brks)
tm_shape(gw_basins_full) +
  tm_polygons(col = "gini_well_depth", palette = "RdPu", title = "Gini Coefficient of Well Depth") 

gw_basins_full%>%
  ggplot(aes(mean_annual_precip, gini_well_capacity))+
  geom_point()+
  theme_bw()
gw_basins_full%>%
  ggplot(aes(mean_annual_precip, gini_well_depth))+
  geom_point()+
  theme_bw()
gw_basins_full%>%
  ggplot(aes(mean_annual_precip, median_well_capacity))+
  geom_point()+
  theme_bw()
gw_basins_full%>%
  ggplot(aes(mean_annual_precip, median_well_depth))+
  geom_point()+
  theme_bw()
```


```{r}
summary(lm(data = gw_basins_full, median_parcel_area ~ mean_annual_precip))
summary(lm(data = gw_basins_full, median_well_depth ~ mean_annual_precip))
summary(lm(data = gw_basins_full, median_well_capacity ~ mean_annual_precip))

summary(lm(data = gw_basins_full, gini_parcel_area ~ mean_annual_precip)) 
summary(lm(data = gw_basins_full, gini_well_depth ~ mean_annual_precip))
summary(lm(data = gw_basins_full, gini_well_capacity ~ mean_annual_precip)) #sig
summary(lm(data = gw_basins_full, gini_land_value ~ mean_annual_precip))
summary(lm(data = gw_basins_full, gini_ag_value ~ mean_annual_precip)) #sig
summary(lm(data = gw_basins_full, gini_improved_value ~ mean_annual_precip)) 

```

### Critical depletion stuff

Options. Relationship between GW basin's critical depletion rating
and: 1. Mean well depth/yield? 2. Water inequality metric 3. Mean land
size/value 4. Land inequality metric 5. Steepness of land-water
relationship (aka, in more critical depleted aquifers, is the
relationship between land ownership and gw access amplified?)

Critically overdrafted subbasins obtained from [CA Dept of Water
Resources](https://water.ca.gov/-/media/DWR-Website/Web-Pages/Programs/Groundwater-Management/Basin-Prioritization/Files/CODBasins_websitemapPAO_a_20y.pdf)

```{r}
overdrafted_subbasins = c("5-022.01", "5-022.04", "5-022.05", "5-022.06", "5-022.07", "5-022.08", "5-022.09", "5-022.11", "5-022.12", "5-022.13", "5-022.14", "6-054")

gw_basins_full = gw_basins_full%>%
  mutate(critically_overdrafted = as.factor(ifelse(subbasin_number %in% overdrafted_subbasins, 1, 0)))

summary(lm(data = gw_basins_full, gini_well_capacity ~ critically_overdrafted))
summary(lm(data = gw_basins_full, gini_well_depth ~ critically_overdrafted))
summary(lm(data = gw_basins_full, gini_improved_value ~ critically_overdrafted))

ggplot(gw_basins_full, aes(critically_overdrafted, gini_well_capacity))+
  geom_boxplot(fill = "coral1")+
  theme_bw()+
  labs(x = "Critically Overdrafted Status", y = "Gini Coefficient of Well Capacity")
ggplot(gw_basins_full, aes(critically_overdrafted, median_well_capacity))+
  geom_boxplot()+
  theme_bw()
ggplot(gw_basins_full, aes(critically_overdrafted, gini_well_depth))+
  geom_boxplot(fill = "skyblue2")+
  theme_bw()+
  labs(x = "Critically Overdrafted Status", y = "Gini Coefficient of Well Depth")
ggplot(gw_basins_full, aes(critically_overdrafted, median_well_depth))+
  geom_boxplot()+
  theme_bw()


t.test(gini_well_capacity ~ critically_overdrafted, data = gw_basins_full)
wilcox.test(gini_well_capacity ~ critically_overdrafted, data = gw_basins_full)

t.test(median_well_depth ~ critically_overdrafted, data = gw_basins_full, var.equal = TRUE)

```

## Crop trends

```{r}
#MAKE SURE TO FILTER OUT PUBLIC LANDS
cropland_parcels_w_crop_mode = master_cropland_parcels%>%
  filter(!owner %in% exclude_owners)%>%
  mutate(exact_extract(cropland_extent_raster, master_cropland_parcels, fun = "mode", progress = TRUE))%>%
  st_join(., gw_basins%>%select(basin_name:subbasin_name), join = st_intersects, largest = T)%>%
  data.frame()%>%
  left_join(., data.frame(levels(cropland_extent_raster)), by = c("exact_extract....." = "Value"))%>%
  filter(!grepl("Developed|Fallow|Wetlands|Pasture|Shrubland|Grass Seed|Open Water|Forest|Barren|Aquaculture", Class_Name))%>%
  mutate(
    crop_group = case_when(
      grepl("Chickpea|Dry Beans|Peas|Vetch|Clover", Class_Name) ~ "Legumes",
      grepl("Safflower|Flaxseed|Canola|Sunflower", Class_Name) ~ "Oil Crops",
      grepl("Apple|Appricot|Avocado|Cantaloupe|Cherries|Citrus|Grapes|Nectarines|Peaches|Pears|Plums|Pomegranate|Olive|Watermelon|Melons", Class_Name) ~ "Fruits",
      grepl("Almond|Walnut|Pecan|Pistachio", Class_Name) ~ "Nuts",
      grepl("Broccoli|Cabbage|Carrot|Cucumber|Garlic|Lettuce|Pepper|Onion|Pumpkin|Squash|Tomato|Turnip|Greens|Herbs", Class_Name) ~ "Vegetables",
      grepl("Barley|Corn|Wheat|Oats|Rice|Rye|Sorghum|Dbl Crop|Triticale", Class_Name) ~ "Grains"),
    crop_group_broad = case_when(
      crop_group %in% c("Legumes", "Grains") | Class_Name == "Alfalfa" ~ "Staple",
      crop_group %in% c("Oil Crops, Fruits, Vegetables, Nuts") | grepl("Misc Vegs|Other Tree Crops|Sugarbeets", Class_Name) ~ "Cash Crop"))


cropland_parcels_w_crop_mode%>%
  group_by(owner, basin_name)%>%
  reframe(parcel_area_ha = sum(parcel_area_ha))%>%
  group_by(crop_group_broad)%>%
  reframe(crop_area_gini = ineq(parcel_area_ha))
cropland_parcels_w_crop_mode%>%
  group_by(crop_group)%>%
  reframe(crop_area_gini = ineq(parcel_area_ha))
cropland_parcels_w_crop_mode%>%
  group_by(Class_Name)%>%
  reframe(crop_area_gini = ineq(parcel_area_ha))

master_cropland_parcels%>%data.frame()%>%
  group_by(owner, )

final_matches_joined_filtered_owners

final_matches_joined_filtered_owners%>%
  mutate(crop = exact_extract(cropland_extent_raster, final_matches_joined_filtered_owners, fun = "mode", progress = TRUE))%>%
  left_join(., data.frame(levels(cropland_extent_raster)), by = c("crop" = "Value"))%>%
  filter(!grepl("Developed|Fallow|Wetlands|Pasture|Shrubland|Grass Seed|Open Water|Forest|Barren|Aquaculture", Class_Name))%>%
  mutate(
    crop_group = case_when(
      grepl("Chickpea|Dry Beans|Peas|Vetch|Clover", Class_Name) ~ "Legumes",
      grepl("Safflower|Flaxseed|Canola|Sunflower", Class_Name) ~ "Oil Crops",
      grepl("Apple|Appricot|Avocado|Cantaloupe|Cherries|Citrus|Grapes|Nectarines|Peaches|Pears|Plums|Pomegranate|Olive|Watermelon|Melons", Class_Name) ~ "Fruits",
      grepl("Almond|Walnut|Pecan|Pistachio", Class_Name) ~ "Nuts",
      grepl("Broccoli|Cabbage|Carrot|Cucumber|Garlic|Lettuce|Pepper|Onion|Pumpkin|Squash|Tomato|Turnip|Greens|Herbs", Class_Name) ~ "Vegetables",
      grepl("Barley|Corn|Wheat|Oats|Rice|Rye|Sorghum|Dbl Crop|Triticale", Class_Name) ~ "Grains"),
    crop_group_broad = case_when(
      crop_group %in% c("Legumes", "Grains") | Class_Name == "Alfalfa" ~ "Staple",
      crop_group %in% c("Oil Crops, Fruits, Vegetables, Nuts") | grepl("Misc Vegs|Other Tree Crops|Sugarbeets", Class_Name) ~ "Cash Crop"))
```

```{r}
# Extract land use values and area fractions for each parcel
land_use_breakdown <- exact_extract(land_use_raster, parcels, include_area = TRUE, progress = TRUE)

# Create an empty data frame to store the results
results <- data.frame()

# Loop through each parcel and process its land use data
for (i in 1:length(land_use_breakdown)) {
  # Get the parcel geometry and its corresponding land use breakdown
  parcel_geom <- parcels[i, ]
  parcel_land_use <- land_use_breakdown[[i]]
  
  # Calculate the total area of the parcel
  total_area <- sum(parcel_land_use$area)
  
  # Create a data frame for each land use within the parcel
  parcel_df <- parcel_land_use %>%
    group_by(value) %>%  # 'value' is the land use category
    summarize(crop_area = sum(area)) %>%  # Calculate the area for each crop
    mutate(crop_percentage = crop_area / total_area,  # Calculate percentage of total area
           parcel_id = parcel_geom$ID,  # Assuming the parcels have an 'ID' column
           geometry = st_geometry(parcel_geom))  # Retain the geometry
  
  # Bind this data to the results dataframe
  results <- bind_rows(results, parcel_df)
}

# Convert to an sf object with the parcel geometries if necessary
results_sf <- st_as_sf(results)
```



### Which crops are most common for farms with most land/water access?

```{r}
hist(master_cropland_parcels$fraction_cropland)
```

### Crop-specific inequality trends/ginis

For land, water, and land-water relationship

Also by subbasin (since crop distribution might differ quite a bit by subbasin)

```{r}

```

### Are bigger farms more likely to grow revenue-intensive or water-intensive crops?

-   USDA crop price data & county-crop-specific yield (Landon has
    dataset)
-   Landon has another dataset of farm costs/budgetings by crop (or
    could even be subsidies)

```{r}

```

## Owner trends

classify owners into groups (ex: LLCs like in Rempel et al) and assess
size by group, crop behaviour by group, etc.